Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-29 WRITEUP/ms.tex
comments on the summary.  Will now try to go back and comment secs 5-7

We show in \Fig{MR} the size-mass relation of the MCCs of \flower compared to observational data of molecular clouds in the Milky Way that are found to be associated with massive \SF \citep{Beuther02a, Mueller02a, Hill05a, Motte07a} and an empirical relation obtained for massive \SF based on Milky Way clouds ($<$10\,\Msun; \citealt{Kauffmann10b, Kauffmann10c} and references therein): $M \gtrsim 870(R/{\rm pc})^{4/3} M_\odot$.
%
In the same plot, we also show loci of constant surface densities, which are parameterized through the visual extinction $A_V$. As shown in \Fig{MR}, the MCCs identified in \flower lie above this relation, along the locus of $A_v \simeq 8.9$ clouds observed by \citet{Lombardi10a} in the MW.
%
%The values of $A_V$ inferred from http://adsabs.harvard.edu/abs/2017MNRAS.471.5018Fthe $M_{\rm cl}$-size relation is consistent with the extinction inferred from gas clumps analysing the UV and IR emission of \flower~by calculating the effect of radiative transfer through dust \citep[][in particular see Fig. 5 therein]{Behrens18a}

\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.8\textwidth]{\figpath/lf16_cloud-mass_size-pc.png}
\caption{
Size-mass relation of MCCs identified in the accreting phase \MM{In
  Figure  of \flower in our simulation (star symbols) compared to observational data of molecular clouds in the Milky Way associated with massive \SF (magenta circles, green stars, blue dots, and black triangles) and empirical relations established based on \obs of the Milky Way. Red line shows the 
mass-size relation found for regions in the Milky Way with massive \SF reported by \citet{Kauffmann10b}. Star symbols are color-coded by increasing $n_{\rm cut}$. Literature data are compiled from \citet{Beuther02a, Mueller02a, Hill05a, Motte07a}. 
Colored lines show the loci expected for different visual extinctions ($A_V$, see \Eq{constantcolumndensity}), which correspond to lines of
constant surface density (i.e., Larson's third relation). This representation is motivated by observational studies (see text and e.g., \citealt{Lombardi10a}).
The star symbols are color-coded by $n_{\rm cut}$, same as \Fig{larsons_single}. 
Extrapolating the masses and sizes of MCCs of \flower along the \citet{Kauffmann10b} 
relation down to 1\,pc are comparable  to those  of  the  cores  found  in regions  of high-mass star formation, suggesting that the identified MCCs are capable of high-mass \SF.
\label{fig:MR}}
\end{figure*}

@@

The largest molecular structure identified is essentially the main disk of \flower and breaks down into smaller substructures that are denser as we increase $n_{\rm cut}$ (see e.g., \Fig{alpha16-28}). Overall, the MCCs we identified are much bigger in size and mass than nearby GMCCs, but comparable to those observed in $z$\,\ssim2$-$4 galaxies based on spatially resolved imaging \citep{Swinbank11a}.

@@

The apparently large virial parameters seen in the MCCs of \flower can be understood using the Toomre $Q$ parameter (see \Fig{Qeff}). We first note that fragmentation {\it can} happen in regions of low $Q$ (if we only consider instability against axisymmetric perturbations), but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$. Second, this fragmentation scale is greater than the typical size of GMCCs. This could be interpreted as a result of instability setting the scales for fragmentation, but the star-forming GMCCs correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales.
%
Thus, the high virial parameter found for most MCCs in \flower, which is the ratio between the velocity dispersion (accounts for effect of pressure and rotation) and surface density, indicate that most of them are not collapsing structures. As such, 
some MCCs are found to lie in regions of $Q_{\rm eff}>1$.
% some MCCs at Q < 1, some are marginally stable against m=0 mode perturb, some are Q>1.
Only in (the denser) regions with a low Toomre $Q_{\rm eff}$ parameter and when the cooling time is lower then the dynamical time star formation can proceed in {\it clumps} and {\it cores} with low virial parameters.

@@

We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by metals, and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping tiles
by identifying a set of different density contours at different evolutionary stages.
Such identification method is similar to MCC identification based on 
contour levels/surfaces of surface brightness maps/cubes of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter for each MCC and we perform a Toomre-$Q$ stability analysis. The properties of the MCCs in \flower are compared with observed MCCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.

We find that the MCC of \flower are highly supersonic, with high velocity dispersions ($\sigma\simeq$\,200\,\kms) comparable to
those observed in $z$\ssim2 starburst galaxies.
The mass scale of the MCCs is of the order of $10^{5.5-8.5}$\,\Msun, depending on the chosen H$_{2}$ cut. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower which break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
The more massive and bigger MCCs in \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which set the scale for fragmentation.
This physical scale is
consistent with what has been found in higher resolution idealized simulation of isolated galaxies.
Our simulation here confirms these results at EoR and under
the influence of continuous gas accretion from the surrounding IGM (see e.g., \citealt{Klessen10a, Goldbaum11a}).

We compare the dynamics of MCCs in \flower to \obs in the context of the Larson's relation.
The MCCs of \flower are found to have higher $\sigma$ and $\Sigma$ systematically regardless of the
density threshold $n_{\rm cut}$ adopted in identifying the structures.
 The found scaling relations are insensitive to the various density cuts. 
 The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the density threshold and 
 even for the molecular substructures. 
 Such a high velocity dispersion likely results from the strong supernova and stellar feedback \flower experienced 
 over the multiple episodes of bursty \SF.

Virial analysis indicates that MCC/arms of the main disk of \flower are unbound, but substructures have
lower virial parameter. Some of these sub-MCCs have Toomre $Q_{\rm eff}$ parameters consistent with being
gravitationally unstable. This is consistent with the notion that collapsing structures resulting from
gravitational instability occur within globally stable structures, which are
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\simeq$\,2 for MCCs in the satellite galaxies, which we interpret as
result of the weaker/delayed stellar feedback compared to \flower. 
In particular, some of the MCCs in the satellites are collapsing structures that have not yet formed stars, as confirmed by their
higher gas-to-stellar fractions.
In terms of the scaling relations examined, we do not observe any significant
temporal variations in the MCC dynamics over the course of 700\,Myr traced in our simulation.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and is subject to the sub-grid models adopted,
serve as an useful avenue/tool for examining and making predictions on the morphology and dynamics of
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.

@@

We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data, and Jens for useful discussions.
%
TKDL acknowledges support by the NSF through award SOSPA4-009 from the NRAO and support from the Simons Foundation.
%
AF acknowledges support from the ERC Advanced Grant INTERSTELLAR H2020/740120.
%
This work is based on a project developed at the Kavli Summer Program in Astrophysics (KSPA) held at the Center for Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli Foundation and the Simons Foundation.
%
We thank the KSPA Scientific and Local Organizing Committees, and the program founder, Pascale Garaud, for supporting the genesis of this work.
%
We also thank the Center for Cosmology and Particle Physics at the New York University for their hospitality in hosting us after the steam pipe explosion in NYC during the KSPA.
%
This research has made use of NASA's Astrophysics Data System Bibliographic Services.
%
We acknowledge use of the Python programming language \citep{VanRossum1991}, Astropy \citep{astropy}, Cython \citep{behnel2010cython}, Matplotlib \citep{Hunter2007}, NumPy \citep{VanDerWalt2011}, \ncode{Pymses} \citep{Labadens2012}, SciPy \citep{scipyref}, and \ncode{yt} \citep{Smith09a,Turk11a}.
%
%and \ncode{yt}, a community-developed Python package for analyzing simulation data in Astronomy.

>>>>

We show in \Fig{MR} the size-mass relation of the MCCs of \flower compared to observational data of molecular clouds in the Milky Way that are found to be associated with massive \SF \citep{Beuther02a, Mueller02a, Hill05a, Motte07a} and an empirical relation obtained for massive \SF based on Milky Way clouds ($<$10\,\Msun; \citealt{Kauffmann10b, Kauffmann10c} and references therein): $M \gtrsim 870(R/{\rm pc})^{4/3} M_\odot$.
%
In the same plot, we also show loci of constant surface densities,
which are parameterized through the visual extinction $A_V$. \MM{Why
  do you go through $A_v$ rather than just using surface densities
  directly?} As shown
in \Fig{MR}, the MCCs identified in \flower lie above this relation,
along the locus of $A_v \simeq 8.9$ clouds observed by
\citet{Lombardi10a} in the MW. \MM{Does the variation in surface
  density seen in Figure 6 correspond to the variation in $A_V$ seen
  here?}
%
%The values of $A_V$ inferred from http://adsabs.harvard.edu/abs/2017MNRAS.471.5018Fthe $M_{\rm cl}$-size relation is consistent with the extinction inferred from gas clumps analysing the UV and IR emission of \flower~by calculating the effect of radiative transfer through dust \citep[][in particular see Fig. 5 therein]{Behrens18a}

\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.8\textwidth]{\figpath/lf16_cloud-mass_size-pc.png}
\caption{
Size-mass relation of MCCs identified in the accreting phase of \flower in our simulation (star symbols) compared to observational data of molecular clouds in the Milky Way associated with massive \SF (magenta circles, green stars, blue dots, and black triangles) and empirical relations established based on \obs of the Milky Way. Red line shows the 
mass-size relation found for regions in the Milky Way with massive \SF
reported by \citet{Kauffmann10b}. Star symbols are color-coded by
increasing $n_{\rm cut}$. \MM{which direction? White high or blue
  high?} Literature data are compiled from \citet{Beuther02a, Mueller02a, Hill05a, Motte07a}. 
Colored lines show the loci expected for different visual extinctions ($A_V$, see \Eq{constantcolumndensity}), which correspond to lines of
constant surface density (i.e., Larson's third relation). This representation is motivated by observational studies (see text and e.g., \citealt{Lombardi10a}).
%mm [redundant] The star symbols are color-coded by $n_{\rm cut}$,
%same as \Fig{larsons_single}.  
Extrapolating the masses and sizes of MCCs of \flower along the \citet{Kauffmann10b} 
relation down to 1\,pc are comparable  to those  of  the  cores  found  in regions  of high-mass star formation, suggesting that the identified MCCs are capable of high-mass \SF.
\label{fig:MR}}
\end{figure*}

@@

The largest molecular structure identified is essentially the main
disk of \flower and breaks down into smaller substructures that are
denser as we increase $n_{\rm cut}$ (see e.g.,
\Fig{alpha16-28}). Overall, the MCCs we identified are much bigger in
size and mass than nearby giant molecular clouds (GMCs), but comparable to those observed in $z$\,\ssim2$-$4 galaxies based on spatially resolved imaging \citep{Swinbank11a}.

@@

The apparently large virial parameters seen in the MCCs of \flower can be understood using the Toomre $Q$ parameter (see \Fig{Qeff}). We first note that fragmentation {\it can} happen in regions of low $Q$ (if we only consider instability against axisymmetric perturbations), but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$. Second, this fragmentation scale is greater than the typical size of GMCs. This could be interpreted as a result of instability setting the scales for fragmentation, but the star-forming GMCs correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales.
%
Thus, the high virial parameter found for most MCCs in \flower, which is the ratio between the velocity dispersion (accounts for effect of pressure and rotation) and surface density, indicate that most of them are not collapsing structures. As such, 
some MCCs are found to lie in regions of $Q_{\rm eff}>1$.
% some MCCs at Q < 1, some are marginally stable against m=0 mode perturb, some are Q>1.
Only in (the denser) regions with a low Toomre $Q_{\rm eff}$ parameter and when the cooling time is lower then the dynamical time star formation can proceed in {\it clumps} and {\it cores} with low virial parameters.

@@

\MM{Each item in the summary needs to refer to definite results and
  analysis in the previous paper.  I have noted the items that don't
  seem to me to have that necessary support.}
We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using the state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by
%mm metals
    UV radiation and metal lines, 
and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping 
%mm tiles
     objects
by identifying a set of 
%mm different density contours at different 
     density contours with varying cutoff densities at multiple
evolutionary stages.
Such an identification method is 
%mm similar 
     the three-dimensional analog
to MCC identification based on 
%mm contour levels/surfaces of surface brightness maps/cubes 
    intensity levels in surface brightness maps or
    position-position-velocity cubes
of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter for each MCC and
we perform a Toomre-$Q$ stability analysis
%mm
    on the galaxy. 
The properties of the MCCs in \flower are compared with observed MCCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.

We find that the MCCs of \flower 
%mm are highly supersonic, with high 
   have highly supersonic
velocity dispersions ($\sigma\simeq$\,200\,\kms) comparable to
those observed in $z$\ssim2 starburst galaxies.
The mass scale of the MCCs is of the order of $10^{5.5-8.5}$\,\Msun,
depending on the chosen H$_{2}$ density cut. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower. These break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
The more massive and bigger MCCs in \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which set the scale for fragmentation. \MM{This could be quantified in
  the discussion by calculating the Jeans scale for the disk conditions.}
This physical scale is
consistent with what has been found in higher resolution idealized
simulation of isolated galaxies. \MM{Perhaps this could be shown more explicitly?}
Our simulation here confirms these results at EoR and under
the influence of continuous gas accretion from the surrounding IGM (see e.g., \citealt{Klessen10a, Goldbaum11a}).

We compare the dynamics of MCCs in \flower to \obs in the context of
the Larson's relation. \MM{Which one?}
The MCCs of \flower are found to have higher $\sigma$ and $\Sigma$
\MM{compared to Milky Way clouds?}
systematically regardless of the 
density threshold $n_{\rm cut}$ adopted in identifying the structures.
The scaling relations found are insensitive to the various density
cuts.  \MM{This point needs to be made more clearly, probably by
  defining the colors that correspond to the density cuts in the figures.}
 The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the density threshold and 
 even for the molecular substructures.  \MM{You never clearly identify
   what you mean by ``molecular substructures.''  Do you just mean
   objects found with high $n_{\rm cut}$?}
 Such a high velocity dispersion likely results from the strong supernova and stellar feedback \flower experienced 
 over the multiple episodes of bursty \SF.   \MM{Or is it from the
   deep potential?  Feedback-driven turbulence famously decays quickly.}

Virial analysis indicates that MCC/arms of the main disk of \flower are unbound, but substructures have
lower virial parameter.
\MM{where is this explicitly discussed?}
 Some of these sub-MCCs have Toomre $Q_{\rm eff}$ parameters consistent with being
gravitationally unstable.  \MM{I don't see this looking at Figure 9.
  Rather the opposite if anything.} This is consistent with the notion that collapsing structures resulting from
gravitational instability occur within globally stable structures, which are
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\simeq$\,2 for MCCs in the satellite
galaxies, \MM{I didn't see any explicit discussion of the MCCs in
  satellite galaxies.} which we interpret as
result of the weaker/delayed stellar feedback compared to \flower. 
In particular, some of the MCCs in the satellites are collapsing structures that have not yet formed stars, as confirmed by their
higher gas-to-stellar fractions. \MM{Also not shown so far as I can see.}
In terms of the scaling relations examined, we do not observe any significant
temporal variations in the MCC dynamics over the course of 700\,Myr
traced in our simulation.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics, and subject to the sub-grid models adopted,
serve as a useful tool for examining and making predictions on the
morphology and dynamics of 
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.

@@

We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data, and Jens for useful discussions.
%
TKDL acknowledges support by the NSF through award SOSPA4-009 from the NRAO and support from the Simons Foundation.
%
AF acknowledges support from the ERC Advanced Grant INTERSTELLAR
H2020/740120.
%mm
    M-MML acknowledges partial support by the NSF through award
    AST18-15461, and hospitality from the Insitute for Theoretical
    Astrophysics at the University of Heidleberg.
%
This work is based on a project developed at the Kavli Summer Program in Astrophysics (KSPA) held at the Center for Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli Foundation and the Simons Foundation.
%
We thank the KSPA Scientific and Local Organizing Committees, and the program founder, Pascale Garaud, for supporting the genesis of this work.
%
We also thank the Center for Cosmology and Particle Physics at the New York University for their hospitality in hosting us after the steam pipe explosion in NYC during the KSPA.
%
This research has made use of NASA's Astrophysics Data System Bibliographic Services.
%
We acknowledge use of the Python programming language \citep{VanRossum1991}, Astropy \citep{astropy}, Cython \citep{behnel2010cython}, Matplotlib \citep{Hunter2007}, NumPy \citep{VanDerWalt2011}, \ncode{Pymses} \citep{Labadens2012}, SciPy \citep{scipyref}, and \ncode{yt} \citep{Smith09a,Turk11a}.
%
%and \ncode{yt}, a community-developed Python package for analyzing simulation data in Astronomy.
